---
title: "Exp1b Ananlysis"
author: "Yiwen Zhang"
date: "2024-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(ggh4x)
library(lmerTest)
library(effsize) 
library(emmeans)
library(patchwork)
library(Hmisc) # for
library(BayesFactor)
```

```{r}
load("/Users/YIZ173/Documents/GitHub/delay_intervening_events_studies/Experiment_1a_1b/exp1b-clean-data.RData")
study0.long <- subj_data %>%
    rename(condition_index = tied_cause,
           causal_strength = value,
           rel_type = relation_type,
           intvn = num_of_intvn)

plot_theme <- theme(
  panel.margin.y = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  text = element_text(size = 16),
  strip.background = element_rect(fill = "white", color = "white"),
  strip.placement = "outside" )

condition_rename <- c(
    "1" = "C1-synced",
    "2" = "C2-synced",
    "3" = "C3-synced"
)
study0.long$condition = condition_rename[as.character(study0.long$condition_index)]
str(study0.long)
```

# Testing for Successful Learning: Veridical Vs Non-veridical

```{r}
# subject level means for veridical and non-veridical ratings
study0.long.subj <- study0.long %>% group_by(task_id, rel_type, condition) %>%
  summarise(cs_value = mean(causal_strength))

# paired t tests
study0.long.subj %>%
  pivot_wider(names_from = rel_type, values_from = cs_value) %>%
  group_by(condition) %>%
  summarise(
    t = t.test(x = veridical, y = nonVeridical, paired = T, alternative = "two.sided")$statistic,
    t.p.value = round(t.test(x = veridical, y = nonVeridical, paired = T, alternative = "two.sided")$p.value,3),
    cohen.d = round( (mean(veridical) - mean(nonVeridical)) / sd(veridical), 3),
    V = wilcox.test(x = veridical, y = nonVeridical, paired = T, alternative = "two.sided",exact = F)$statistic,
    wilcox.p = round(wilcox.test(x = veridical, y = nonVeridical, paired = T, alternative = "two.sided",exact = F)$p.value,3),
    ttestBF = round( exp(ttestBF(x = veridical, y = nonVeridical, paired = T)@bayesFactor$bf), 3)
  ) %>%
mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p, method="BH"))


```

# calculate difference score

```{r}
study0.long.diff.subj <- study0.long %>% 
  group_by(condition, task_id, rel_type, intvn, cause, delay) %>%
  summarise(cs_value = mean(causal_strength)) %>%
  pivot_wider(names_from = rel_type, values_from = cs_value) %>%
  mutate(diff = veridical - nonVeridical) %>%
    ungroup()

study0.long.diff.subj %>%
  group_by(condition, cause, intvn, delay) %>%
  summarise(
    mean = t.test(diff, mu = 0, alternative = "two.sided")$estimate,
    lower = t.test(diff, mu = 0, alternative = "two.sided")$conf.int[1],
    upper = t.test(diff, mu = 0, alternative = "two.sided")$conf.int[2],
    # t.p.value = round(t.test(diff, mu = 0, alternative = "two.sided")$p.value,3),
    # wilcox.p = wilcox.test(diff, mu = 0, alternative = "two.sided",exact = F)$p.value,
    # ttestBF = round( exp(ttestBF(diff, mu = 0)@bayesFactor$bf), 3)
    ) -> diff_smry
```


# spearman correlation 

## testing the impact of intervening events

```{r}
study0.long.diff.subj %>%
    subset(delay %in% c(4,6,8)) %>%
  group_by(delay) %>%
  summarise(
    spearman_rho = cor(intvn, diff, method = "spearman"),
    p_value = cor.test(intvn, diff, method = "spearman")$p.value,
    bayes_factor = exp(correlationBF(rank(intvn), rank(diff))@bayesFactor$bf)
  ) %>%
mutate(p_adjust = p.adjust(p_value, method="BH"))
```

```{r}
# spearman correlation
study0.long.diff.subj %>%
  group_by(intvn) %>%
  summarise(
    spearman_rho = cor(delay, diff, method = "spearman"),
    p_value = cor.test(delay, diff, method = "spearman")$p.value,
    bayes_factor = exp(correlationBF(rank(delay), rank(diff))@bayesFactor$bf)
  ) -> delay_spearman
delay_spearman %>%
mutate(p_adjust = p.adjust(p_value, method="BH"))
```

# Figures

## Figure for the impact of intervening events


```{r fig.height=4, fig.width=6}

study0.long %>% 
  group_by(condition, task_id, rel_type, intvn, cause, delay) %>%
    summarise(cs_value = mean(causal_strength)) %>% # get subject level mean
    group_by(condition, rel_type, intvn, cause, delay) %>%
  summarise(
    mean = t.test(cs_value, mu = 0, alternative = "two.sided")$estimate,
    lower = t.test(cs_value, mu = 0, alternative = "two.sided")$conf.int[1],
    upper = t.test(cs_value, mu = 0, alternative = "two.sided")$conf.int[2],
    ) -> rel_type_bycond_smry

rel_type_bycond_smry$rel_type = factor(rel_type_bycond_smry$rel_type, levels = c("veridical", "nonVeridical"), label = c("Veridical", "Non Veridical"))

rel_type_bycond_smry %>%
    mutate(cause_name = ifelse(intvn > 0 , paste0(toupper(cause), "\n(",intvn, " ICs)"), paste0(toupper(cause), "\n(",intvn, " IC)"))) %>%
    subset(delay %in% c('4', '6', '8')) %>%
    mutate(delay_name = paste("Delay =", delay, "hours")) %>%
  ggplot(aes(x = factor(intvn), y = mean, color = rel_type, group = rel_type)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper), width = 0.3) +
  theme_bw() +
  plot_theme +
  scale_color_manual(values = c("lightblue","#1E5162")) +
  theme(legend.position = "bottom") +
    ylim(0,10) +
  facet_grid2(~delay_name, scales = "free") +
  labs(x = "# of Intervening Causes", y = "Causal Strength", color = "") -> p1.1

p1.1

# ggsave("../figures/study1_long_diff_score_intvn.png",
#   plot = p1,
#   scale = 3,
#   width = 800,
#   height = 300,
#   units = c("px"),
#   dpi = 300)
```

## Figure for the impact of delays

```{r fig.height=4, fig.width=6}

rel_type_bycond_smry %>%
    mutate(intvn_name = ifelse(intvn > 1, paste(intvn,"ICs"), paste(intvn,"IC"))) %>%
  ggplot(aes(x = factor(delay), y = mean, color = rel_type, group = rel_type)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper), width = 0.3) +
  theme_bw() +
  plot_theme +
  scale_color_manual(values = c("lightblue","#1E5162")) +
  theme(legend.position = "bottom") +
    ylim(0,10) +
  facet_grid2(~intvn_name, scales = "free") +
  labs(x = "Delay (hours)", y = "Causal Strength", color = "") -> p2.1

p2.1

```

# Compare Exp 1a and 1b

```{r}
# prepare Exp 1a data
study0.short1 <- read.csv("/Users/YIZ173/Documents/GitHub/delay_intervening_events_studies/Experiment_1a_1b/exp1a-clean-data.csv")
# calculate different score
study0.short.diff.subj <- study0.short1 %>% 
  group_by(dataset, task_id, rel_type, intvn, cause_num) %>%
  summarise(cs_value = mean(causal_strength)) %>%
  pivot_wider(names_from = rel_type, values_from = cs_value) %>%
  mutate(diff = veridical - nonVeridical)

short_diff <- study0.short.diff.subj %>%
   subset(dataset == "three_pos_10_2_0_12.csv") %>%
    group_by(task_id) %>%
    summarise(mean_diff = mean(diff)) %>%
    mutate(exp = "Experiment 1a\n(short timeframe)")
long_diff <- study0.long.diff.subj %>%
    group_by(task_id) %>%
    summarise(mean_diff = mean(diff)) %>%
    mutate(exp = "Experiment 1b\n(long timeframe)")

comp_timeframe <- rbind(short_diff, long_diff)

comp_timeframe %>%
    group_by(exp) %>%
    summarise(
        mean = mean(mean_diff),
        lower = t.test(mean_diff, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[1],
        upper = t.test(mean_diff, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[2]) %>%
  ggplot(aes(x = exp, y = mean, color = exp)) +
  geom_col(fill = "white", size = 2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.3) +
  theme_bw() +
  plot_theme +
  theme(legend.position = "none") +
  labs(x = "", y = "Difference Score")-> p4
p4

```

```{r}
t.test(mean_diff~exp, data = comp_timeframe, alternative = "two.sided")
wilcox.test(mean_diff~exp, data = comp_timeframe, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)
cohen.d(mean_diff~exp, data = comp_timeframe)
ttestBF(x = short_diff$mean_diff, y = long_diff$mean_diff)
```



# Other analyses in registration (not included in the manuscript)

## t tests for the impact of intervening events

```{r}
# Define a helper function to run the statistical tests and return the results
run_tests <- function(data, delay_value, intvn_levels, comparison_label, exp_label) {
  data %>%
    subset(delay == delay_value & intvn %in% intvn_levels) %>%
    mutate(intvn = factor(intvn)) %>%
    summarise(
      t = round(t.test(diff ~ intvn, alternative = "two.sided")$statistic,3),
      t.p.value = round(t.test(diff ~ intvn, alternative = "two.sided")$p.value,3),
      cohen.d = round(cohen.d(diff ~ intvn, na.rm = TRUE)$estimate,3),
      W = round(wilcox.test(diff ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,3),
      wilcox = round(wilcox.test(diff ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,3),
      BF = round(exp(ttestBF(formula = diff ~ intvn, data = cur_data())@bayesFactor$bf), 3)
    ) %>%
    mutate(
      comparison = comparison_label,
      delay = paste0(delay_value, " hrs"),
      exp = exp_label
    )
}

# List of comparisons and delays
comparisons <- list(
  list(delay_value = '4', intvn_levels = c('0', '1'), comparison = "0 vs 1 event", exp = "long"),
  list(delay_value = '8', intvn_levels = c('1', '2'), comparison = "1 vs 2 events", exp = "long"),
  list(delay_value = '6', intvn_levels = c('0', '1'), comparison = "0 vs 1 event", exp = "long"),
  list(delay_value = '6', intvn_levels = c('1', '2'), comparison = "1 vs 2 events", exp = "long"),
  list(delay_value = '6', intvn_levels = c('0', '2'), comparison = "0 vs 2 events", exp = "long")
)

# Run the tests for each comparison and combine the results
results <- lapply(comparisons, function(comp) {
  run_tests(
    data = study0.long.diff.subj,
    delay_value = comp$delay_value,
    intvn_levels = comp$intvn_levels,
    comparison_label = comp$comparison,
    exp_label = comp$exp
  )
})

# Combine all results into one data frame
final_results <- do.call(rbind, results)

# View the final combined results
final_results$t.p.adj <- p.adjust(final_results$t.p.value, method = "fdr")
final_results$wilcox.p.adj <- p.adjust(final_results$wilcox, method = "fdr")
  
# View the final combined results with adjusted p-values
final_results %>% arrange(delay, comparison)

```



## t tests for the impact of delays


```{r}
# Define a helper function to run the statistical tests and return the results
run_tests <- function(data, intvn_value, delay_values, comparison_label, exp_label) {
  data %>%
    subset(intvn == intvn_value & delay %in% delay_values) %>%
    mutate(intvn = factor(intvn)) %>%
    summarise(
      t = round(t.test(diff ~ delay, alternative = "two.sided")$statistic,3),
      t.p.value = round(t.test(diff ~ delay, alternative = "two.sided")$p.value,3),
      cohen.d = round(cohen.d(diff ~ delay, na.rm = TRUE)$estimate,3),
      W = round(wilcox.test(diff ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,3),
      wilcox = round(wilcox.test(diff ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,3),
      BF = round(exp(ttestBF(formula = diff ~ delay, data = cur_data())@bayesFactor$bf), 3)
    ) %>%
    mutate(
      comparison = comparison_label,
      intvn = intvn_value,
      exp = exp_label
    )
}

# List of comparisons and corresponding parameters
comparisons <- list(
  list(intvn_value = '0', delay_values = c('2', '4'), comparison = "2 vs 4", exp = "long"),
  list(intvn_value = '0', delay_values = c('4', '6'), comparison = "4 vs 6", exp = "long"),
  list(intvn_value = '0', delay_values = c('2', '6'), comparison = "2 vs 6", exp = "long"),
  list(intvn_value = '1', delay_values = c('4', '6'), comparison = "4 vs 6", exp = "long"),
  list(intvn_value = '1', delay_values = c('6', '8'), comparison = "6 vs 8", exp = "long"),
  list(intvn_value = '1', delay_values = c('4', '8'), comparison = "4 vs 8", exp = "long"),
  list(intvn_value = '2', delay_values = c('6', '8'), comparison = "6 vs 8", exp = "long"),
  list(intvn_value = '2', delay_values = c('8', '10'), comparison = "8 vs 10", exp = "long"),
  list(intvn_value = '2', delay_values = c('6', '10'), comparison = "6 vs 10", exp = "long")
)

# Run the tests for each comparison and combine the results
results <- lapply(comparisons, function(comp) {
  run_tests(
    data = study0.long.diff.subj,
    intvn_value = comp$intvn_value,
    delay_values = comp$delay_values,
    comparison_label = comp$comparison,
    exp_label = comp$exp
  )
})

# Combine all results into one data frame
final_results <- do.call(rbind, results)

# View the final combined results
final_results$t.p.adj <- p.adjust(final_results$t.p.value, method = "fdr")
final_results$wilcox.p.adj <- p.adjust(final_results$wilcox, method = "fdr")
  
# View the final combined results with adjusted p-values
final_results %>% arrange(intvn, comparison)


```



## Testing for Veridical ratings only

### the influence of intervening events

```{r fig.height=3, fig.width=8}
study0.long %>%
  subset(rel_type == "veridical") -> veri_df.long


veri_df.long$delay = factor(veri_df.long$delay)
veri_df.long$intvn = factor(veri_df.long$intvn)

veri_df.long %>%
  # subset(delay %in% c(3,9)) %>%
  group_by(delay, intvn, condition_index) %>%
  summarise(
    mean = mean(causal_strength),
    lower = t.test(causal_strength, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[1],
    upper = t.test(causal_strength, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[2],
    # t.p.value = t.test(causal_strength, mu = 0, alternative = "two.sided")$p.value,
    # wilcox.p = wilcox.test(causal_strength, mu = 0, alternative = "two.sided",exact = F)$p.value,
    # cohen.d = cohen.d(causal_strength, f=NA, mu = 0,na.rm =T)$estimate,
    # BF =exp(ttestBF(causal_strength, mu = 0)@bayesFactor$bf)
  ) -> veri_df_smry


# the effect of intervening events
veri_df_smry %>%
    subset(delay %in% c('4', '6', '8')) %>%
    mutate(delay_name = paste("delay =", delay, "hrs")) %>%
  ggplot(aes(x = intvn, y = mean, group = intvn)) +
  geom_col(position=position_dodge(width=0.3), fill = "lightgrey") +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_grid(~delay_name) +
  labs(y = "Causal Strength \n Veridical", x = "# of intervening events") +
  theme_bw() +
  plot_theme +
theme(legend.position = "none") 
  # ylim(0,10) +
  # scale_fill_hue(h = c(180, 300)) 
```

```{r}
veri_df.long %>%
  subset(delay == '6' & intvn %in% c('0', '1')) %>%
mutate(intvn = factor(intvn)) %>%
  summarise(t.p.value = t.test(causal_strength ~ intvn, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ intvn, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ intvn, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "0 vs 1 event", delay = "6 hrs", exp="exp1b")-> intvn_0vs1.long

veri_df.long %>%
  subset(delay == '6' & intvn %in% c('1', '2')) %>%
    mutate(intvn = factor(intvn)) %>%
  summarise(t.p.value = t.test(causal_strength ~ intvn, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ intvn,na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ intvn, data = cur_data())@bayesFactor$bf),3)
            ) %>%
  mutate(comparison = "1 vs 2 events", delay = "6 hrs", exp="exp1b")-> intvn_1vs2.long

veri_df.long %>%
  subset(delay == '6' & intvn %in% c('0', '2')) %>%
    mutate(intvn = factor(intvn)) %>%
  summarise(t.p.value = t.test(causal_strength ~ intvn, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ intvn,na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ intvn, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "0 vs 2 events", delay = "6 hrs", exp="exp1b")-> intvn_0vs2.long

rbind(intvn_0vs1.long,
      intvn_1vs2.long,
      intvn_0vs2.long)

```


### the influence of delay

```{r fig.height=3, fig.width=8, paged.print=FALSE}
veri_df_smry %>%
mutate(intvn_name = paste(intvn,"intervening causes")) %>%
  ggplot(aes(x = delay, y = mean, group = delay)) +
  # geom_hline(yintercept = 0, alpha = 0.5) +
  geom_col(fill = "lightblue") +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_wrap(~intvn_name, scales = "free_x") +
  labs(y = "Causal Strength \n Veridical", x = "Delay (hrs)") +
  theme_bw() +
  plot_theme +
theme(legend.position = "none") +
  # ylim(0,10) +
  scale_color_hue(h = c(180, 300)) 
```

```{r}
veri_df.long %>%
  subset(intvn == '0' & delay %in% c('2', '4')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "2 vs 4", intvn = 0, exp="exp1b")-> delay_2vs4_0e.long

veri_df.long %>%
  subset(intvn == '0' & delay %in% c('4', '6')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "4 vs 6", intvn = 0, exp="exp1b")-> delay_4vs6_0e.long

veri_df.long %>%
  subset(intvn == '1' & delay %in% c('4', '6')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "4 vs 6", intvn = 1, exp="exp1b")-> delay_4vs6_1e.long

veri_df.long %>%
  subset(intvn == '1' & delay %in% c('8', '6')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "6 vs 8", intvn = 1, exp="exp1b")-> delay_6vs8_1e.long

veri_df.long %>%
  subset(intvn == '2' & delay %in% c('6', '8')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "6 vs 8", intvn = 2, exp="exp1b")-> delay_6vs8_2e.long

veri_df.long %>%
  subset(intvn == '2' & delay %in% c('8', '10')) %>%
mutate(delay = factor(delay)) %>%
  summarise(t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(causal_strength ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "8 vs 10", intvn = 2, exp="exp1b")-> delay_10vs8_2e.long


rbind(delay_2vs4_0e.long,
      delay_4vs6_0e.long,
      delay_4vs6_1e.long,
      delay_6vs8_1e.long,
      delay_6vs8_2e.long,
      delay_10vs8_2e.long
)
```


## Testing for Nonverdicals

```{r}
# visualizations
study0.long %>%
  subset(rel_type == "nonVeridical") %>%
    group_by(task_id, delay, intvn, condition_index) %>%
    summarise(cs = mean(causal_strength))-> nonVeri_df.long


nonVeri_df.long$delay = factor(nonVeri_df.long$delay)
nonVeri_df.long$intvn = factor(nonVeri_df.long$intvn)

nonVeri_df.long %>%
  # subset(delay %in% c(3,9)) %>%
  group_by(delay, intvn, condition_index) %>%
  summarise(
    mean = mean(cs),
    lower = t.test(cs, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[1],
    upper = t.test(cs, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[2],
    # t.p.value = t.test(causal_strength, mu = 0, alternative = "two.sided")$p.value,
    # wilcox.p = wilcox.test(causal_strength, mu = 0, alternative = "two.sided",exact = F)$p.value,
    # cohen.d = cohen.d(causal_strength, f=NA, mu = 0,na.rm =T)$estimate,
    # BF =exp(ttestBF(causal_strength, mu = 0)@bayesFactor$bf)
  ) -> nonVeri_df_smry


# the effect of intervening events
nonVeri_df_smry %>%
    subset(delay == '6') %>%
  ggplot(aes(x = intvn, y = mean, color = intvn, group = intvn)) +
  # geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position=position_dodge(width=0.3)) +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  # facet_grid(~delay) +
  labs(y = "Causal Strength \n Non-veridical", x = "# of intervening events") +
  theme_bw() +
  plot_theme +
theme(legend.position = "none") +
  ylim(0,10) +
  scale_color_hue(h = c(180, 300)) 
```
```{r}
nonVeri_df_smry %>%
  ggplot(aes(x = delay, y = mean, color = delay, group = delay)) +
  # geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position=position_dodge(width=0.3)) +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_wrap(~intvn, scales = "free_x") +
  labs(y = "Causal Strength \n Non-veridical", x = "Delay (hrs)") +
  theme_bw() +
  plot_theme +
theme(legend.position = "none") +
  ylim(0,10) +
  scale_color_hue(h = c(180, 300)) 
```

```{r}
t.test(cs ~ intvn, alternative = "two.sided", data = subset(nonVeri_df.long, delay == '6' & intvn %in% c('0', '1')))

nonVeri_df.long %>%
  subset(delay == '6' & intvn %in% c('0', '1')) %>%
  group_by(delay) %>%
  dplyr::summarise(t.p.value = t.test(cs ~ intvn, alternative = "two.sided", data = .)$p.value,
            wilcox = wilcox.test(cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95, data = .)$p.value,
            cohen.d = cohen.d(cs ~ intvn, na.rm =T, data = .)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ intvn, data = .)@bayesFactor$bf),3)) %>%
  mutate(comparison = "0 vs 1 event", delay = "6 hrs", exp="exp1b")
```
### the influence of intervening events

```{r}
nonVeri_df.long %>%
  subset(delay == '6' & intvn %in% c('0', '1')) %>%
  mutate(intvn = factor(intvn)) %>%
  group_by(delay) %>%
  dplyr::summarise(t.p.value = t.test(cs ~ intvn, alternative = "two.sided", data = .)$p.value,
            wilcox = wilcox.test(cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95, data = .)$p.value,
            cohen.d = cohen.d(cs ~ intvn, na.rm =T, data = .)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ intvn, data = .)@bayesFactor$bf),3)) %>%
  mutate(comparison = "0 vs 1 event", delay = "6 hrs", exp="exp1b")-> intvn_0vs1.long.nonVeri

nonVeri_df.long %>%
  subset(delay == '6' & intvn %in% c('1', '2')) %>%
    mutate(intvn = factor(intvn)) %>%
  group_by(delay) %>%
  summarise(t.p.value = t.test(cs ~ intvn, alternative = "two.sided", data = .)$p.value,
            wilcox = wilcox.test(cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95, data = .)$p.value,
            cohen.d = cohen.d(cs ~ intvn,na.rm =T, data = .)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ intvn, data = .)@bayesFactor$bf),3)
            ) %>%
  mutate(comparison = "1 vs 2 events", delay = "6 hrs", exp="exp1b")-> intvn_1vs2.long.nonVeri

nonVeri_df.long %>%
  subset(delay == '6' & intvn %in% c('0', '2')) %>%
    mutate(intvn = factor(intvn)) %>%
  group_by(delay) %>%
  summarise(t.p.value = t.test(cs ~ intvn, alternative = "two.sided", data = .)$p.value,
            wilcox = wilcox.test(cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95, data = .)$p.value,
            cohen.d = cohen.d(cs ~ intvn,na.rm =T, data = .)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ intvn, data = .)@bayesFactor$bf),3)) %>%
  mutate(comparison = "0 vs 2 events", delay = "6 hrs", exp="exp1b")-> intvn_0vs2.long.nonVeri

rbind(intvn_0vs1.long.nonVeri,
      intvn_1vs2.long.nonVeri,
      intvn_0vs2.long.nonVeri)

```

```{r}
nonVeri_df.long %>%
  subset(intvn == '0' & delay %in% c('2', '4')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "2 vs 4", intvn = 0, exp="exp1b")-> delay_2vs4_0e.long.nonVeri

nonVeri_df.long %>%
  subset(intvn == '0' & delay %in% c('4', '6')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "4 vs 6", intvn = 0, exp="exp1b")-> delay_4vs6_0e.long.nonVeri

nonVeri_df.long %>%
  subset(intvn == '1' & delay %in% c('4', '6')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "4 vs 6", intvn = 1, exp="exp1b")-> delay_4vs6_1e.long.nonVeri

nonVeri_df.long %>%
  subset(intvn == '1' & delay %in% c('8', '6')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "6 vs 8", intvn = 1, exp="exp1b")-> delay_6vs8_1e.long.nonVeri

nonVeri_df.long %>%
  subset(intvn == '2' & delay %in% c('6', '8')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "6 vs 8", intvn = 2, exp="exp1b")-> delay_6vs8_2e.long.nonVeri

nonVeri_df.long %>%
  subset(intvn == '2' & delay %in% c('8', '10')) %>%
mutate(delay = factor(delay)) %>%
  group_by(intvn) %>%
  summarise(t.p.value = t.test(cs ~ delay, alternative = "two.sided")$p.value,
            wilcox = wilcox.test(cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(cs ~ delay, na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = cs ~ delay, data = cur_data())@bayesFactor$bf),3)) %>%
  mutate(comparison = "8 vs 10", intvn = 2, exp="exp1b")-> delay_10vs8_2e.long.nonVeri


rbind(delay_2vs4_0e.long.nonVeri,
      delay_4vs6_0e.long.nonVeri,
      delay_4vs6_1e.long.nonVeri,
      delay_6vs8_1e.long.nonVeri,
      delay_6vs8_2e.long.nonVeri,
      delay_10vs8_2e.long.nonVeri
)
```



## descriptive analysis for memory

```{r}
# a histogram look at how many people did the memory question correct
study0.long %>%
    select(task_id, condition_index, mem_cause_1, mem_cause_2, mem_cause_3) %>%
    distinct() %>%
    pivot_longer(mem_cause_1 : mem_cause_3,names_to = "mem_cause", values_to = "cause_remembered") %>%
    na.omit() -> cause_recall

cause_recall %>%
    group_by(cause_remembered, mem_cause) %>%
    summarise(n = n()) %>%
    ggplot(aes(x=factor(cause_remembered), y = n))+
    geom_col() +
    facet_grid(~mem_cause) +
    theme_bw() +
    plot_theme
```

```{r}
cause_recall %>%
    mutate(prompt_cause_num = extract_numeric(mem_cause),
           accuracy = ifelse(prompt_cause_num == cause_remembered, 1, 0)) -> cause_recall_accuracy

cause_recall_accuracy %>%
    group_by(task_id, condition_index) %>%
    summarise(mean_acc = mean(accuracy)) %>%
    subset(mean_acc > 0) -> good_memory_list

cause_recall_accuracy %>%
    group_by(task_id) %>%
    summarise(mean_acc = mean(accuracy)) %>%
    count(mean_acc)  %>%
    mutate(percentage = n/298)
    # group_by(condition_index) %>%
    # summarise(mean = mean(accuracy),
    #           lower = t.test(accuracy, mu = 0, alternative = "two.sided")$conf.int[1],
    # upper = t.test(accuracy, mu = 0, alternative = "two.sided")$conf.int[2])
    
    
```

```{r}
good_memory_list %>%
    ungroup() %>%
    count(condition_index)
```


