---
title: "Experiment 2a and 2b Analysis"
author: "Yiwen Zhang"
date: "2024-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggh4x)
library(gridExtra)
library(lme4)
library(lmerTest)
library(effsize) 
library(emmeans)
library(Hmisc) # for
library(BayesFactor)
library(brms)
library(bayestestR)
```


#------veridical agian zero------
## Causal Strength

```{r}
load("/Users/YIZ173/Documents/GitHub/delay_intervening_events_studies/Experiment_2a_2b/exp2-clean-data2.RData")
cs_df.short$exp = "Experiment 2a \n (short timeframe)"
cs_df.long$exp = "Experiment 2b \n (long timeframe)"

cs_df <- rbind(cs_df.short, cs_df.long) %>%
  mutate(condition = paste("Condition", condition_index))

plot_theme <- theme(
  panel.margin.y = unit(2, "lines"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  text = element_text(size = 16),
  strip.background = element_rect(fill = "white", color = "white"),
  strip.placement = "outside",
  legend.position = "bottom"
)
```



```{r cars, fig.height=6, fig.width=10}
# causal strength against 0
cs_df %>%
  group_by(exp, user_id, rel_type, cs_type, condition) %>%
  summarise(cs_value = mean(causal_strength)) -> cs_subj_df

cs_subj_df %>%
  group_by(exp, rel_type, cs_type, condition) %>%
  summarise(
    mean = t.test(cs_value, mu = 0, alternative = "two.sided")$estimate,
    lower = t.test(cs_value, mu = 0, alternative = "two.sided")$conf.int[1],
    upper = t.test(cs_value, mu = 0, alternative = "two.sided")$conf.int[2],
    t = round(t.test(cs_value, mu = 0, alternative = "two.sided")$statistic,3),
    t.p.value = t.test(cs_value, mu = 0, alternative = "two.sided")$p.value,
    cohen.d = round(cohen.d(cs_value, f=NA, mu = 0,na.rm =T)$estimate,3),
    wilcox = round(wilcox.test(cs_value, mu = 0, alternative = "two.sided",exact = F)$statistic,3),
    wilcox.p = wilcox.test(cs_value, mu = 0, alternative = "two.sided",exact = F)$p.value,
    BF =exp(ttestBF(cs_value, mu = 0)@bayesFactor$bf)) -> rel_type_smry
rel_type_smry$cs_type = factor(rel_type_smry$cs_type, levels = c('6', '12', '18', '24'))
```


## Predictive Strength

```{r}
pred_p_df.short$exp = "Experiment 2a \n (short timeframe)"
pred_p_df.long$exp = "Experiment 2b \n (long timeframe)"

pred_p_df <- rbind(pred_p_df.short, pred_p_df.long) %>%
  mutate(condition = paste("Condition", condition_index))
pred_p_df$half <- factor(pred_p_df$half, levels = c("1-6", "7-12", "13-18", "19-24"))
pred_p_df$delay <- factor(pred_p_df$delay)

pred_p_df %>%
  group_by(exp, condition, half) %>%
  summarise(
    mean = t.test(pred_strength, mu = 0, alternative = "two.sided")$estimate,
    lower = t.test(pred_strength, mu = 0, alternative = "two.sided")$conf.int[1],
    upper = t.test(pred_strength, mu = 0, alternative = "two.sided")$conf.int[2],
    t = round(t.test(pred_strength, mu = 0, alternative = "two.sided")$statistic,3),
    t.p.value = t.test(pred_strength, mu = 0, alternative = "two.sided")$p.value,
    cohen.d = round(cohen.d(pred_strength, f=NA, mu = 0,na.rm =T)$estimate,3),
    wilcox = round(wilcox.test(pred_strength, mu = 0, alternative = "two.sided",exact = F)$statistic,3),
    wilcox.p = wilcox.test(pred_strength, mu = 0, alternative = "two.sided",exact = F)$p.value,
    BF =exp(ttestBF(pred_strength, mu = 0)@bayesFactor$bf)) -> pred_strength_q1


# pred_strength_q1$t.p.adjust = p.adjust(pred_strength_q1$t.p.value, method="BH")
# pred_strength_q1$wilcox.p.adjust = p.adjust(pred_strength_q1$wilcox.p, method="BH")
pred_strength_smry <- pred_strength_q1 %>%
  arrange(exp, condition, half) %>%
  select(exp, condition, half, mean, lower, upper, t, t.p.value, cohen.d, wilcox, wilcox.p, BF) # %>%
  # write.csv(file = "tables/study2_veriNonVeri_comparison_pred_strength.csv")

pred_strength_smry
```
## Adjust p values for causal strength and predictive strength together

if cs_type == "N/A", the rows were for preditive strength
if half == "N/A", the rows were for causal strength

```{r}
# adjust p value
rel_type_test <- rel_type_smry %>%
  ungroup() %>%
  subset(rel_type == "veridical") %>%
    arrange(exp, condition, cs_type) %>%
  select(exp, condition, cs_type, mean, lower, upper, t, t.p.value, cohen.d, wilcox, wilcox.p, BF)

rbind(pred_strength_smry, rel_type_test) %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p, method="BH")) %>%
  select(exp, condition, cs_type, half, mean, lower, upper, t, t.p.value, t.p.adjust, cohen.d, wilcox, wilcox.p, wilcox.p.adjust, BF) ->  veri_against_0_results_bind
# veri_against_0_results_bind %>%  write.csv(file = "tables/study2_veri_against_0.csv")

veri_against_0_results_bind
```

## figures


```{r fig.height=6, fig.width=10}
# figure for causal strength against 0 in exp 2a and 2b
veri_against_0_results_bind %>%
  subset(!is.na(cs_type)) %>%
  ggplot(aes(x = factor(cs_type), y = mean, group = 1)) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_line(linetype="dashed", position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper),width = 0.5) +
  facet_grid2(rows = vars(exp), cols = vars(condition), axes = "x", switch = "y") +
  scale_y_continuous(breaks = seq(-1,4,1), limits = c(-0.3,4.3)) +
  scale_color_manual(values = c("black","#162a33")) +
  theme_bw() +
  plot_theme +
  theme(
        legend.position = "none",
        legend.direction="vertical")+
  labs(x = "Trial", y = "Causal Strength", color = "", shape = "") +
  geom_text(aes(label = ifelse(t.p.adjust < 0.001, "***", ifelse( t.p.adjust < 0.01, "**", ifelse(t.p.adjust < 0.05, "*", "")))),
    y = 4, vjust=0, size = 5)-> p1
p1
# ggsave(plot = p1, filename = "figures/Study 2 veridical causal against 0.png",
#        scale = 3,
#         width = 1000,
#         height = 500,
#         units = c("px"),
#         dpi = 300)
```

```{r fig.height=6, fig.width=10}
# figure for predictive strength against 0 in exp 2a and 2b
veri_against_0_results_bind %>%
  subset(is.na(cs_type)) %>%
  ggplot(aes(x = factor(half), y = mean, group = 1)) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_line(position = position_dodge(width = 0.5), linetype = "dashed") + 
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper), width = 0.5) +
  facet_grid2(rows = vars(exp), cols = vars(condition), axes = "x", switch = "y") +
  ylab("Predictive Strength") +
  labs(x = "Trial", y = "Predictive Strength", color = "", shape = "") +
  scale_y_continuous(breaks = seq(-1,4,1), limits = c(-0.3,4.3)) +
  theme_bw() +
  plot_theme  +
  theme(
        legend.position = "bottom",
        legend.direction="vertical") +
  geom_text(aes(label = ifelse(t.p.adjust < 0.001, "***", ifelse(t.p.adjust < 0.01, "**", ifelse(t.p.adjust < 0.05, "*", "")))),
    y = 4, vjust=0, size = 5)-> p4
p4
# ggsave(plot = p4, filename = "figures/Study 2 prediction strength agianst 0.png",
#        scale = 3,
#         width = 1000,
#         height = 500,
#         units = c("px"),
#         dpi = 300)
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


#-----intervening events------

## prep data

```{r}
# prep data and code the conditions for causal strength
conditions.short <- 
  data.frame(condition_index = c(1,1,2,2,3,3),
             cause_order = c("c1", "c3", "c1", "c3", "c1", "c3"),
             # effect_order = c("e1","e3","e1","e3","e1","e3"),
             delay = c(3, 1, 9, 3, 11, 9),
             intvn = c("2 events","no event","2 events","no event","2 events","no event"))

cs_df.short %>%
  subset(rel_type == "veridical") %>%
  left_join(conditions.short) -> veri_df.short


# condition for exp 2b
conditions.long <- 
  data.frame(condition_index = c(1,1,2,2,3,3,4,4),
             cause_order = c("c1", "c3", "c1", "c3", "c1", "c3", "c1", "c3"),
             # effect_order = c("e1","e3","e1","e3","e1","e3"),
             delay = c(3, 1, 9, 3, 11, 9, 0, 0),
             intvn = c("2 events","no event","2 events","no event","2 events","no event","2 events","no event"))

cs_df.long %>%
  subset(rel_type == "veridical") %>%
  left_join(conditions.long) -> veri_df.long

veri_df <- rbind(veri_df.long, veri_df.short)
veri_df$delay = factor(veri_df$delay)
veri_df$intvn = factor(veri_df$intvn)
```

## t tests

```{r warning=TRUE}
## t tests for Causal Strength
veri_df %>%
  subset(delay %in% c(3,9)) %>%
  group_by(exp, cs_type, delay) %>%
  summarise(t = t.test(causal_strength ~ intvn, alternative = "two.sided")$statistic,
    t.p.value = t.test(causal_strength ~ intvn, alternative = "two.sided")$p.value,
            V = wilcox.test(causal_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
            wilcox.p = wilcox.test(causal_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = round(cohen.d(causal_strength ~ intvn,na.rm =T)$estimate,3),
            BF = round(exp(ttestBF(formula = causal_strength ~ intvn, data = cur_data())@bayesFactor$bf),3)) %>%
  arrange(exp, delay, cs_type)-> intvn_effects_smry

intvn_effects_smry %>%
  select(exp, cs_type, delay, t, t.p.value, cohen.d, BF, V, wilcox.p) -> intvn_cs_smry

## t tests for predictive Strength
pred_p_df %>%
  subset(delay %in% c(3,9)) %>%
  group_by(exp, half, delay) %>%
  summarise(t = t.test(pred_strength ~ intvn, alternative = "two.sided")$statistic,
    t.p.value = t.test(pred_strength ~ intvn, alternative = "two.sided")$p.value,
    V = wilcox.test(pred_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
            wilcox.p = wilcox.test(pred_strength ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = cohen.d(pred_strength ~ intvn,na.rm =T)$estimate,
            BF =round(exp(ttestBF(formula = pred_strength ~ intvn, data = cur_data())@bayesFactor$bf),3)) -> ps_intvn_effects_smry
  
ps_intvn_effects_smry %>%
  arrange(exp, delay, half) %>%
  select(exp, half, delay, t, t.p.value, cohen.d, BF, V, wilcox.p) -> ps_intvn_effects_smry

# adjusting p values
rbind(intvn_cs_smry, ps_intvn_effects_smry) %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p, method="BH")) %>%
  select(exp, delay, cs_type, half, t, t.p.value, t.p.adjust, cohen.d, V, wilcox.p, wilcox.p.adjust, BF) #%>%
  # write.csv("tables/study2_intvn_comparison.csv")

```


## figures

```{r echo=FALSE, fig.height=6, fig.width=7}
# Causal Strength separated by intervening events
veri_df %>%
  # subset(delay %in% c(3,9)) %>%
  group_by(exp, cs_type, delay, intvn, condition_index) %>%
  summarise(
    mean = mean(causal_strength),
    lower = t.test(causal_strength, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[1],
    upper = t.test(causal_strength, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[2],
    t.p.value = t.test(causal_strength, mu = 0, alternative = "two.sided")$p.value,
    wilcox.p = wilcox.test(causal_strength, mu = 0, alternative = "two.sided",exact = F)$p.value,
    cohen.d = cohen.d(causal_strength, f=NA, mu = 0,na.rm =T)$estimate,
    BF =exp(ttestBF(causal_strength, mu = 0)@bayesFactor$bf)
  ) -> veri_df_smry


# the effect of intervening events
veri_df_smry %>%
  subset(delay %in% c(3,9)) %>%
  mutate(delay_name = paste("Delay =", delay)) %>%
  ggplot(aes(x = factor(cs_type), y = mean, color = intvn, group = intvn)) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position=position_dodge(width=0.3), aes(shape = intvn), size = 3) +
  geom_line(linetype="dashed") +
  # geom_col(position=position_dodge(width=0.9), aes(fill = intvn)) +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_grid2(exp~delay_name, axes = "all") +
  labs(y = "Causal Strength (Veridical)", color = "", shape = "") +
  theme_bw() +
  plot_theme +
  scale_y_continuous(breaks = seq(-1,3,1), limits = c(-1,3)) +
  scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(1, "lines"),
        legend.position = "left",
        legend.direction="vertical") + 
  xlab("Trial") -> p2
p2

# ggsave(plot = p2, filename = "figures/Study 2 the influence of intervening events.png",
#        scale = 3,
#         width = 700,
#         height = 600,
#         units = c("px"),
#         dpi = 300)
```


```{r echo=FALSE, fig.height=6, fig.width=7}
# Predictive Strength separated by intervening events
pred_p_df %>%
  group_by(exp, delay, intvn, half) %>%
  summarise(
    mean = t.test(pred_strength, mu = 0, alternative = "two.sided")$estimate,
    lower = t.test(pred_strength, mu = 0, alternative = "two.sided")$conf.int[1],
    upper = t.test(pred_strength, mu = 0, alternative = "two.sided")$conf.int[2],
    t.p.value = t.test(pred_strength, mu = 0, alternative = "two.sided")$p.value,
    cohen.d = cohen.d(pred_strength, f=NA, mu = 0,na.rm =T)$estimate,
    BF =exp(ttestBF(pred_strength, mu = 0)@bayesFactor$bf)) %>%
  arrange(half, intvn, delay)-> pred_strength_df_smry

pred_strength_df_smry %>%
  subset(delay %in% c(3,9)) %>%
  ggplot(aes(x = factor(half), y = mean, color = intvn, group = intvn)) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position=position_dodge(width=0.3), aes(shape = intvn), size = 3) +
  geom_line(linetype="dashed") +
  # geom_col(position=position_dodge(width=0.9), aes(fill = intvn)) +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_grid2(exp~delay, axes = "all") +
  labs(y = "Predictive Strength", color = "", shape = "") +
  theme_bw() +
  plot_theme +
  # scale_y_continuous(breaks = seq(-1,3,1), limits = c(-1,3)) +
  scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(1, "lines"),
        legend.position = "left",
        legend.direction="vertical") + 
  xlab("Trial") ->  p5
p5

# ggsave(plot = p5, filename = "figures/Study 2 the influence of intervening events (pred strength).png",
#        scale = 3,
#         width = 700,
#         height = 600,
#         units = c("px"),
#         dpi = 300)
```



#-----the influence of delay------

## spearman correlation tests

```{r}
# spearman correlation for causal strength
veri_df %>%
  mutate(delay_num = as.numeric(as.character(delay))) %>%
  group_by(exp, cs_type, intvn) %>%
  summarise(
    spearman_rho = cor(delay_num, causal_strength, method = "spearman"),
    p_value = cor.test(delay_num, causal_strength, method = "spearman")$p.value,
    bayes_factor = exp(correlationBF(rank(delay_num), rank(causal_strength))@bayesFactor$bf)
  ) -> delay_spearman

# spearman correlation for causal strength
pred_p_df %>%
  mutate(delay_num = as.numeric(as.character(delay))) %>%
  group_by(exp, half, intvn) %>%
  summarise(
    spearman_rho = cor(delay_num, pred_strength, method = "spearman"),
    p_value = cor.test(delay_num, pred_strength, method = "spearman")$p.value,
    bayes_factor = exp(correlationBF(rank(delay_num), rank(pred_strength))@bayesFactor$bf)
  ) -> delay_spearman_pred

# adjusting p values
delay_spearman$measure = "causal_strength"
delay_spearman_pred$measure = "pred_strength"
rbind(delay_spearman, delay_spearman_pred) %>%
  group_by(exp) %>%
  mutate(p.adjust = p.adjust(p_value, method="BH")) %>%
  arrange(exp, measure, intvn, cs_type) %>%
  select(exp,measure, intvn, cs_type,spearman_rho,p_value,p.adjust,bayes_factor)  -> delay_spearman_combined
delay_spearman_combined
# delay_spearman_combined %>%
#   write.csv("tables/study2_delay_spearman.csv")
```

## figures

```{r}
# figure for causal strength separated by delay
veri_df_smry %>%
  # subset(exp == "Experiment 2a \n (short timeframe)") %>%
  ggplot(aes(x = cs_type, y = mean, color = factor(delay), group = factor(delay)))+
  # geom_line(position=position_dodge(width=0.5), aes(linetype = factor(delay))) +
  geom_line(position=position_dodge(width=0.5), linetype = "dotted") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin = lower, ymax = upper), width = 0.3) +
  geom_point(position=position_dodge(width=0.5), size = 6, fill = "white", shape = 21) +
  geom_text(aes(label = factor(delay)), position=position_dodge(width=0.5) ) +
  facet_nested(exp~intvn, axes = "x", scales = "free") +
  ylab("Causal Strength") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  scale_y_continuous(breaks = seq(0,3,1), limits = c(-0.5,3)) +
  scale_color_manual(values = c("#c0c5ce", "#a7adba", "#65737e", "#4f5b66", "#343d46")) + # #c0c5ce
  # scale_colour_brewer(palette = "Set1") +
  # scale_linetype_manual(values = c("solid", "longdash", "dashed", "twodash", "dotted")) +
  xlab("Delay (hours)") +
  plot_theme +
  # scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(0.5, "lines")) -> p3.3
p3.3

# ggsave(plot = p3.3, filename = "figures/Study 2 the influence of delays [color].png",
#        scale = 3,
#         width = 700,
#         height = 700,
#         units = c("px"),
#         dpi = 300)
```
```{r}
# figure for predictive strength separated by delay
pred_strength_df_smry %>%
  # subset(exp == "Experiment 2a \n (short timeframe)") %>%
  ggplot(aes(x = half, y = mean, color = factor(delay), group = factor(delay)))+
  # geom_line(position=position_dodge(width=0.5), aes(linetype = factor(delay))) +
  geom_line(position=position_dodge(width=0.5), linetype = "dotted") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin = lower, ymax = upper), width = 0.3) +
  geom_point(position=position_dodge(width=0.5), size = 6, fill = "white", shape = 21) +
  geom_text(aes(label = factor(delay)), position=position_dodge(width=0.5) ) +
  facet_nested(exp~intvn, axes = "x", scales = "free") +
  ylab("Causal Strength") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  scale_y_continuous(breaks = seq(-1,4,1), limits = c(-0.5,4.5)) +
  scale_color_manual(values = c("#c0c5ce", "#a7adba", "#65737e", "#4f5b66", "#343d46")) + # #c0c5ce
  # scale_linetype_manual(values = c("solid", "longdash", "dashed", "twodash", "dotted")) +
  xlab("Delay (hours)") +
  plot_theme +
  # scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(0.5, "lines")) -> p6.3
p6.3
# ggsave(plot = p6.3, filename = "figures/Study 2 the influence of delays (Pred strength).png",
#        scale = 3,
#         width = 1000,
#         height = 700,
#         units = c("px"),
#         dpi = 300)
```


#-----Comparing conditions-----



```{r fig.height=6, fig.width=5}
# these figures are not included in manuscript but very useful to compare across conditions
rel_type_smry %>%
  subset(rel_type == "veridical") %>%
  ggplot(aes(x = factor(cs_type), y = mean, color = factor(condition), group = factor(condition))) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_line(position = position_dodge(width = 0.5)) +
  # geom_col(position = position_dodge(width = 0.9)) + 
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_wrap(~exp,ncol=1, scales = "free_x") +
  theme_bw() +
  plot_theme +
  theme(legend.position = "left")+
  labs(x = "trial", y = "Causal Strength", color = "Condition")-> p1
p1
```

### exp 2a

#### causal strength

```{r}
# Data preparation
cs_df.short_agg <- cs_df.short %>% 
  subset(rel_type == "veridical") %>% 
  group_by(user_id, condition_index) %>% 
  summarise(cs = mean(causal_strength), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

# Fit model
cs_df.short_agg_model <- lm(cs ~ condition_index, data = cs_df.short_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(cs_df.short_agg_model)
bf_anova <- anovaBF(cs ~ condition_index, data = cs_df.short_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(cs_df.short_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- cs_df.short_agg %>%
  group_by(condition_index) %>%
  summarise(mean_cs = mean(cs), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(cs_df.short_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- cs_df.short_agg$cs[cs_df.short_agg$condition_index == cond1]
    x2 <- cs_df.short_agg$cs[cs_df.short_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```

#### predictive strength

```{r}

# Data preparation
pred_p_df.short_agg <- pred_p_df.short.1_24 %>% 
  group_by(user_id, condition_index) %>% 
  summarise(ps = mean(pred_strength), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

pred_p_df.short_agg_model <- lm(ps ~ condition_index, data = pred_p_df.short_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(pred_p_df.short_agg_model)
bf_anova <- anovaBF(ps ~ condition_index, data = pred_p_df.short_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(pred_p_df.short_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- pred_p_df.short_agg %>%
  group_by(condition_index) %>%
  summarise(mean_ps = mean(ps), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(pred_p_df.short_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- pred_p_df.short_agg$ps[pred_p_df.short_agg$condition_index == cond1]
    x2 <- pred_p_df.short_agg$ps[pred_p_df.short_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```

### exp 2b

#### causal strength

```{r}
# Data preparation
cs_df.long_agg <- cs_df.long %>% 
  subset(rel_type == "veridical") %>% 
  group_by(user_id, condition_index) %>% 
  summarise(cs = mean(causal_strength), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

# Fit model
cs_df.long_agg_model <- lm(cs ~ condition_index, data = cs_df.long_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(cs_df.long_agg_model)
bf_anova <- anovaBF(cs ~ condition_index, data = cs_df.long_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(cs_df.long_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- cs_df.long_agg %>%
  group_by(condition_index) %>%
  summarise(mean_cs = mean(cs), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(cs_df.long_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- cs_df.long_agg$cs[cs_df.long_agg$condition_index == cond1]
    x2 <- cs_df.long_agg$cs[cs_df.long_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```

#### predictive strength

```{r}

# Data preparation
pred_p_df.long_agg <- pred_p_df.long.1_24 %>% 
  group_by(user_id, condition_index) %>% 
  summarise(ps = mean(pred_strength), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

pred_p_df.long_agg_model <- lm(ps ~ condition_index, data = pred_p_df.long_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(pred_p_df.long_agg_model)
bf_anova <- anovaBF(ps ~ condition_index, data = pred_p_df.long_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(pred_p_df.long_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- pred_p_df.long_agg %>%
  group_by(condition_index) %>%
  summarise(mean_ps = mean(ps), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(pred_p_df.long_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- pred_p_df.long_agg$ps[pred_p_df.long_agg$condition_index == cond1]
    x2 <- pred_p_df.long_agg$ps[pred_p_df.long_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```


#-------Exp2a vs Exp2b--------


## t test for aggregated casual strength of short and long

```{r}
cs_df.agg <- rbind(mutate(cs_df.short_agg, exp = "exp 2a"), mutate(cs_df.long_agg, exp = "exp 2b")) %>% mutate(exp = factor(exp))

# Run t-test
t_result <- t.test(cs ~ exp, data = cs_df.agg, alternative = "two.sided")
print(t_result)

# Calculate Cohen's d
cohens_d <- cohen.d(cs ~ exp, data = cs_df.agg)
print(cohens_d)

# Bayes Factor
bf_result <- ttestBF(formula = cs ~ exp, data = cs_df.agg, alternative = "two.sided")
print(bf_result)
```



## t test for aggregated predictive strength of short and long

```{r}
ps_df.agg <- rbind(mutate(pred_p_df.short_agg, exp = "exp 2a"), mutate(pred_p_df.long_agg, exp = "exp 2b")) %>% mutate(exp = factor(exp))

# Run t-test
t_result <- t.test(ps ~ exp, data = ps_df.agg, alternative = "two.sided")
print(t_result)

# Calculate Cohen's d
cohens_d <- cohen.d(ps ~ exp, data = ps_df.agg)
print(cohens_d)

# Bayes Factor
bf_result <- ttestBF(formula = ps ~ exp, data = ps_df.agg, alternative = "two.sided")
print(bf_result)
```


# ------- Appendix or Supplymental Material------------------


## t tests for the effect of delays 

(this is in the registration, but replaced by spearman test in the manuscript)

```{r}
# pairwise t tests for causal strength
run_delay_tests <- function(data, intvn_value, delay_values, comparison_label, exp_label) {
  data %>%
    subset(intvn == intvn_value & delay %in% delay_values) %>%
    mutate(delay = factor(delay, levels = delay_values)) %>%
    group_by(cs_type) %>%
    summarise(
      t.statistic = t.test(causal_strength ~ delay, alternative = "two.sided")$statistic,
      t.p.value = t.test(causal_strength ~ delay, alternative = "two.sided")$p.value,
      wilcox.statistic = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
      wilcox.p.value = wilcox.test(causal_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
      cohen.d = cohen.d(causal_strength ~ delay, na.rm = TRUE)$estimate,
      BF = round(exp(ttestBF(formula = causal_strength ~ delay, data = cur_data())@bayesFactor$bf), 3)
    ) %>%
    mutate(
      comparison = comparison_label,
      intvn = intvn_value,
      exp = exp_label
    )
}

# List of comparisons and corresponding parameters
comparisons <- list(
  list(data = veri_df.long, intvn_value = "no event", delay_values = c(0, 1), comparison = "0 vs 1", exp = "long"),
  list(data = veri_df.long, intvn_value = "no event", delay_values = c(1, 3), comparison = "1 vs 3", exp = "long"),
  list(data = veri_df.long, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = veri_df.long, intvn_value = "2 events", delay_values = c(0, 3), comparison = "0 vs 3", exp = "long"),
  list(data = veri_df.long, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = veri_df.long, intvn_value = "2 events", delay_values = c(9, 11), comparison = "9 vs 11", exp = "long"),
  list(data = veri_df.short, intvn_value = "no event", delay_values = c(1, 3), comparison = "1 vs 3", exp = "short"),
  list(data = veri_df.short, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = veri_df.short, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = veri_df.short, intvn_value = "2 events", delay_values = c(9, 11), comparison = "9 vs 11", exp = "short")
)

# Run the tests for each comparison and combine the results
results <- lapply(comparisons, function(comp) {
  run_delay_tests(
    data = comp$data,
    intvn_value = comp$intvn_value,
    delay_values = comp$delay_values,
    comparison_label = comp$comparison,
    exp_label = comp$exp
  )
})

# Combine all results into one data frame
delay_comparisons <- do.call(rbind, results)

# View the final combined results
delay_comparisons

```


```{r}
# pairwise t tests for predictive strength
run_delay_tests <- function(data, intvn_value, delay_values, comparison_label, exp_label) {
  data %>%
    subset(intvn == intvn_value & delay %in% delay_values) %>%
    mutate(delay = factor(delay, levels = delay_values)) %>%
    group_by(half) %>%
    summarise(
      t.statistic = t.test(pred_strength ~ delay, alternative = "two.sided")$statistic,
      t.p.value = t.test(pred_strength ~ delay, alternative = "two.sided")$p.value,
      wilcox.statistic = wilcox.test(pred_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
      wilcox.p.value = wilcox.test(pred_strength ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
      cohen.d = cohen.d(pred_strength ~ delay, na.rm = TRUE)$estimate,
      BF = round(exp(ttestBF(formula = pred_strength ~ delay, data = cur_data())@bayesFactor$bf), 3)
    ) %>%
    mutate(
      comparison = comparison_label,
      intvn = intvn_value,
      exp = exp_label
    )
}

# List of comparisons and corresponding parameters
comparisons <- list(
  list(data = pred_p_df.long, intvn_value = "no event", delay_values = c(0, 1), comparison = "0 vs 1", exp = "long"),
  list(data = pred_p_df.long, intvn_value = "no event", delay_values = c(1, 3), comparison = "1 vs 3", exp = "long"),
  list(data = pred_p_df.long, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = pred_p_df.long, intvn_value = "2 events", delay_values = c(0, 3), comparison = "0 vs 3", exp = "long"),
  list(data = pred_p_df.long, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = pred_p_df.long, intvn_value = "2 events", delay_values = c(9, 11), comparison = "9 vs 11", exp = "long"),
  list(data = pred_p_df.short, intvn_value = "no event", delay_values = c(1, 3), comparison = "1 vs 3", exp = "short"),
  list(data = pred_p_df.short, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = pred_p_df.short, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = pred_p_df.short, intvn_value = "2 events", delay_values = c(9, 11), comparison = "9 vs 11", exp = "short")
)

# Run the tests for each comparison and combine the results
results <- lapply(comparisons, function(comp) {
  run_delay_tests(
    data = comp$data,
    intvn_value = comp$intvn_value,
    delay_values = comp$delay_values,
    comparison_label = comp$comparison,
    exp_label = comp$exp
  )
})

# Combine all results into one data frame
delay_comparisons_ps <- do.call(rbind, results) %>% rename(cs_type = half)

# View the final combined results
delay_comparisons_ps 

```


```{r}
## adjust p values
delay_comparisons$measure = "causal_strength"
delay_comparisons_ps$measure = "pred_strength"
rbind(delay_comparisons, delay_comparisons_ps) %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p.value, method="BH")) %>%
  select(exp, measure, intvn, cs_type, comparison, t.statistic, t.p.value, t.p.adjust, cohen.d, wilcox.statistic, wilcox.p.value, wilcox.p.adjust, BF) %>%
  arrange(exp, measure, intvn, cs_type, comparison) %>%
  write.csv("tables/study2_delay_comparison.csv")
```



#-----Analyses for Non Veridical-----



```{r}
## figures for raw ratings against 0
rel_type_smry$cs_type = factor(rel_type_smry$cs_type, levels = c('6', '12', '18', '24'))
rel_type_smry %>%
  subset(rel_type == "nonVeridical") %>%
  ggplot(aes(x = factor(cs_type), y = mean, color = rel_type, group = rel_type)) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  geom_line(linetype="dashed", position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = rel_type)) +
  geom_errorbar(position = position_dodge(width = 0.5), aes(ymin = lower, ymax = upper),width = 0.5) +
  facet_grid2(rows = vars(exp), cols = vars(condition), axes = "all", switch = "y") +
  scale_y_continuous(breaks = seq(-1,3,1), limits = c(-1,3)) +
  scale_color_manual(values = c("black","#162a33")) +
  theme_bw() +
  plot_theme +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(1, "lines"),
        legend.position = "none",
        legend.direction="none")+
  labs(x = "Trial", y = "Causal Strength (Non-veridical)", color = "", shape = "")-> p1
p1
# ggsave(plot = p1, filename = "figures/Study 2 Non-veridical causal against 0.png",
#        scale = 3,
#         width = 1000,
#         height = 500,
#         units = c("px"))

# test the raw rating against 0
rel_type_smry %>%
  subset(rel_type == "nonVeridical") %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p, method="BH")) %>%
  select(exp, condition, cs_type, t, t.p.value, t.p.adjust, cohen.d, wilcox, wilcox.p, wilcox.p.adjust, BF) %>%
  arrange(exp, condition, cs_type) # %>%
  # write.csv(file = "tables/study2_non-veri_against_0.csv")
```


## getting absolute values for the non-veridical data

```{r}
cs_df.long %>%
  subset(rel_type == "nonVeridical") %>%
  subset(cause_order != "c2") %>%
  left_join(conditions.long) %>% mutate(abs_cs = abs(causal_strength)) -> nonVeri_abs.long

cs_df.short %>%
  subset(rel_type == "nonVeridical") %>%
  subset(cause_order != "c2") %>%
  left_join(conditions.short) %>% mutate(abs_cs = abs(causal_strength))-> nonVeri_abs.short

nonVeri_abs <- rbind(nonVeri_abs.long, nonVeri_abs.short) %>% mutate(abs_cs = abs(causal_strength))

nonVeri_abs %>%
  # subset(delay %in% c(3,9)) %>%
  group_by(exp, cs_type, delay, intvn, condition_index) %>%
  summarise(
    mean = mean(abs_cs),
    lower = t.test(abs_cs, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[1],
    upper = t.test(abs_cs, mu = 0, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$conf.int[2],
    t.p.value = t.test(abs_cs, mu = 0, alternative = "two.sided")$p.value,
    wilcox.p = wilcox.test(abs_cs, mu = 0, alternative = "two.sided",exact = F)$p.value,
    cohen.d = cohen.d(abs_cs, f=NA, mu = 0,na.rm =T)$estimate,
    BF =exp(ttestBF(abs_cs, mu = 0)@bayesFactor$bf)
  ) -> nonVeri_abs_smry
nonVeri_abs_smry
```

## intervening events

```{r}
# t test for the impact of intervening events
nonVeri_abs %>%
  subset(delay %in% c(3,9)) %>%
  mutate(intvn = factor(intvn)) %>%
  group_by(exp, cs_type, delay) %>%
  summarise(t = t.test(abs_cs ~ intvn, alternative = "two.sided")$statistic,
    t.p.value = t.test(abs_cs ~ intvn, alternative = "two.sided")$p.value,
            V = wilcox.test(abs_cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
            wilcox.p = wilcox.test(abs_cs ~ intvn, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
            cohen.d = round(cohen.d(abs_cs ~ intvn,na.rm =T)$estimate,3),
            BF = round(exp(ttestBF(formula = abs_cs ~ intvn, data = cur_data())@bayesFactor$bf),3)) %>%
  arrange(exp, delay, cs_type)-> nonveri_intvn_effects_smry


nonveri_intvn_effects_smry %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p, method="BH")) %>%
  select(exp, delay, cs_type, t, t.p.value, t.p.adjust, cohen.d, V, wilcox.p, wilcox.p.adjust, BF) %>%
   write.csv("tables/study2_non-veri_intvn_comparison.csv")
```


```{r}
# the figure of intervening events
nonVeri_abs_smry %>%
  subset(delay %in% c(3,9)) %>%
  mutate(delay_name = paste("Delay =", delay)) %>%
  ggplot(aes(x = factor(cs_type), y = mean, color = intvn, group = intvn)) +
  # geom_hline(yintercept = 0, alpha = 0.5) +
  geom_point(position=position_dodge(width=0.3), aes(shape = intvn), size = 3) +
  geom_line(linetype="dashed") +
  # geom_col(position=position_dodge(width=0.9), aes(fill = intvn)) +
  geom_errorbar(position=position_dodge(width=0.3), aes(ymin = lower, ymax = upper), width = 0.3) +
  facet_grid2(exp~delay_name, axes = "all") +
  labs(y = "Causal Strength (Veridical)", color = "", shape = "") +
  theme_bw() +
  plot_theme +
  scale_y_continuous(breaks = seq(0,3,1), limits = c(0,3)) +
  scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(1, "lines"),
        legend.position = "none") + 
  xlab("Trial") -> p2
p2

ggsave(plot = p2, filename = "figures/Study 2 Non-veridical - intervening events.png",
       scale = 3,
        width = 700,
        height = 500,
        units = c("px"),
        dpi = 300)
```


## delay

```{r}
# spearman t test
nonVeri_abs %>%
  mutate(delay_num = as.numeric(as.character(delay))) %>%
  group_by(exp, cs_type, intvn) %>%
  summarise(
    spearman_rho = cor(delay_num, causal_strength, method = "spearman"),
    p_value = cor.test(delay_num, causal_strength, method = "spearman")$p.value,
    bayes_factor = exp(correlationBF(rank(delay_num), rank(causal_strength))@bayesFactor$bf)
  ) %>%
  mutate(p.adjust = p.adjust(p_value, method="BH")) %>%
  arrange(exp, intvn, cs_type) %>%
  select(exp, intvn, cs_type,spearman_rho,p_value,p.adjust,bayes_factor)-> delay_spearman_nonVeri

delay_spearman_nonVeri

# delay_spearman_nonVeri %>%
#   write.csv("tables/study2_non-veridical_delay_spearman.csv")
```

```{r fig.height=5, fig.width=8}
nonVeri_abs_smry %>%
  # subset(exp == "Experiment 2a \n (short timeframe)") %>%
  ggplot(aes(x = cs_type, y = mean, color = factor(delay), group = factor(delay)))+
  # geom_line(position=position_dodge(width=0.5), aes(linetype = factor(delay))) +
  geom_line(position=position_dodge(width=0.6), linetype = "dotted") +
  # geom_hline(yintercept = 0, alpha = 0.5) +
  geom_errorbar(position=position_dodge(width=0.6), aes(ymin = lower, ymax = upper), width = 0.3) +
  geom_point(position=position_dodge(width=0.6), size = 6, fill = "white", shape = 21) +
  geom_text(aes(label = factor(delay)), position=position_dodge(width=0.6) ) +
  facet_nested(exp~intvn, axes = "x", scales = "free") +
  ylab("Causal Strength") +
  theme_bw() +
  scale_y_continuous(breaks = seq(0,3,1), limits = c(0,3)) +
  scale_color_manual(values = c("#c0c5ce", "#a7adba", "#65737e", "#4f5b66", "#343d46")) + # #c0c5ce
  # scale_linetype_manual(values = c("solid", "longdash", "dashed", "twodash", "dotted")) +
  xlab("Delay (hours)") +
  plot_theme +
  # scale_color_hue(h = c(180, 300)) +
  theme(panel.margin.y = unit(3, "lines"),
        panel.margin.x = unit(0.5, "lines"),
        legend.position = "none") -> p3.5
p3.5
# ggsave(plot = p3.3, filename = "figures/Study 2 Non-veridical - delays.png",
#        scale = 3,
#         width = 700,
#         height = 500,
#         units = c("px"),
#         dpi = 300)
```


```{r}
# t tests for the impact of delay (registered but not in the manuscript)

run_non_veri_delay_tests <- function(data, intvn_value, delay_values, comparison_label, exp_label) {
  data %>%
    subset(intvn == intvn_value & delay %in% delay_values) %>%
    mutate(delay = factor(delay)) %>%
    group_by(cs_type) %>%
    summarise(
      t.statistic = t.test(abs_cs ~ delay, alternative = "two.sided")$statistic,
      t.p.value = t.test(abs_cs ~ delay, alternative = "two.sided")$p.value,
      wilcox.statistic = wilcox.test(abs_cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$statistic,
      wilcox.p.value = wilcox.test(abs_cs ~ delay, alternative = "two.sided", conf.int = TRUE, conf.level = 0.95)$p.value,
      cohen.d = cohen.d(abs_cs ~ delay, na.rm = TRUE)$estimate,
      BF = round(exp(ttestBF(formula = abs_cs ~ delay, data = cur_data())@bayesFactor$bf), 3)
    ) %>%
    mutate(
      comparison = comparison_label,
      intvn = intvn_value,
      exp = exp_label
    )
}

# List of comparisons and corresponding parameters
comparisons <- list(
  list(data = nonVeri_abs.long, intvn_value = "no event", delay_values = c(0, 1), comparison = "0 vs 1", exp = "long"),
  list(data = nonVeri_abs.long, intvn_value = "no event", delay_values = c(3, 1), comparison = "1 vs 3", exp = "long"),
  list(data = nonVeri_abs.long, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = nonVeri_abs.long, intvn_value = "2 events", delay_values = c(0, 3), comparison = "0 vs 3", exp = "long"),
  list(data = nonVeri_abs.long, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "long"),
  list(data = nonVeri_abs.long, intvn_value = "2 events", delay_values = c(11, 9), comparison = "9 vs 11", exp = "long"),
  list(data = nonVeri_abs.short, intvn_value = "no event", delay_values = c(3, 1), comparison = "1 vs 3", exp = "short"),
  list(data = nonVeri_abs.short, intvn_value = "no event", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = nonVeri_abs.short, intvn_value = "2 events", delay_values = c(3, 9), comparison = "3 vs 9", exp = "short"),
  list(data = nonVeri_abs.short, intvn_value = "2 events", delay_values = c(11, 9), comparison = "9 vs 11", exp = "short")
)

# Run the tests for each comparison and combine the results
results <- lapply(comparisons, function(comp) {
  run_non_veri_delay_tests(
    data = comp$data,
    intvn_value = comp$intvn_value,
    delay_values = comp$delay_values,
    comparison_label = comp$comparison,
    exp_label = comp$exp
  )
})

# Combine all results into one data frame
delay_comparisons <- do.call(rbind, results)

# View the final combined results
delay_comparisons %>%
  group_by(exp) %>%
  mutate(t.p.adjust = p.adjust(t.p.value, method="BH"),
         wilcox.p.adjust = p.adjust(wilcox.p.value, method="BH")) %>%
  select(exp, intvn, cs_type, comparison, t.statistic, t.p.value, t.p.adjust, cohen.d, wilcox.statistic, wilcox.p.value, wilcox.p.adjust, BF) %>%
  arrange(exp, intvn, cs_type, comparison) %>%
  write.csv("tables/study2_non-veridical_delay_comparison.csv")

```

## comparing conditions

```{r}
# Data preparation
nonVeri_abs.short_agg <- nonVeri_abs %>% 
  subset(rel_type == "nonVeridical" & exp == "Experiment 2a \n (short timeframe)") %>% 
  group_by(user_id, condition_index) %>% 
  summarise(cs.abs = mean(abs_cs), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

# Fit model
nonVeri_abs.short_agg_model <- lm(cs.abs ~ condition_index, data = nonVeri_abs.short_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(nonVeri_abs.short_agg_model)
bf_anova <- anovaBF(cs.abs ~ condition_index, data = nonVeri_abs.short_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(nonVeri_abs.short_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- nonVeri_abs.short_agg %>%
  group_by(condition_index) %>%
  summarise(mean_cs_abs = mean(cs.abs), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(nonVeri_abs.short_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- nonVeri_abs.short_agg$cs.abs[nonVeri_abs.short_agg$condition_index == cond1]
    x2 <- nonVeri_abs.short_agg$cs.abs[nonVeri_abs.short_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```

```{r}
# Data preparation
nonVeri_abs.long_agg <- nonVeri_abs %>% 
  subset(rel_type == "nonVeridical" & exp == "Experiment 2b \n (long timeframe)") %>% 
  group_by(user_id, condition_index) %>% 
  summarise(cs.abs = mean(abs_cs), .groups = 'drop') %>% 
  mutate(condition_index = factor(condition_index))

# Fit model
nonVeri_abs.long_agg_model <- lm(cs.abs ~ condition_index, data = nonVeri_abs.long_agg)

# Create results table
library(knitr)

# 1. ANOVA Results Table
cat("ANOVA RESULTS\n")
cat("=============\n")
anova_results <- joint_tests(nonVeri_abs.long_agg_model)
bf_anova <- anovaBF(cs.abs ~ condition_index, data = nonVeri_abs.long_agg)

anova_table <- data.frame(
  Test = "Condition Effect",
  F_value = round(anova_results$F.ratio, 3),
  df = paste(anova_results$df1, anova_results$df2, sep = ", "),
  p_value = ifelse(anova_results$p.value < 0.001, "<0.001", round(anova_results$p.value, 3)),
  BF10 = round(exp(bf_anova@bayesFactor$bf), 3)
)

print(anova_table, row.names = FALSE)

# 2. Pairwise Comparisons Table
cat("\n\nPAIRWISE COMPARISONS\n")
cat("====================\n")

# Get pairwise results
post_hoc <- emmeans(nonVeri_abs.long_agg_model, pairwise ~ condition_index, adjust = "BH")
pairwise_results <- summary(post_hoc$contrasts)

# Get condition means for effect sizes
condition_means <- nonVeri_abs.long_agg %>%
  group_by(condition_index) %>%
  summarise(mean_cs_abs = mean(cs.abs), .groups = 'drop')

# Calculate Bayes Factors for each comparison
conditions <- levels(nonVeri_abs.long_agg$condition_index)
pairwise_table <- data.frame()

for (i in 1:(length(conditions)-1)) {
  for (j in (i+1):length(conditions)) {
    cond1 <- conditions[i]
    cond2 <- conditions[j]
    
    # Get data for each condition
    x1 <- nonVeri_abs.long_agg$cs.abs[nonVeri_abs.long_agg$condition_index == cond1]
    x2 <- nonVeri_abs.long_agg$cs.abs[nonVeri_abs.long_agg$condition_index == cond2]
    
    # Calculate BF
    bf_result <- ttestBF(x = x1, y = x2)
    
    # Find corresponding row in pairwise results
    contrast_name <- paste(cond1, "-", cond2)
    pairwise_row <- pairwise_results[grepl(paste(cond1, ".*", cond2, sep=""), pairwise_results$contrast) | 
                                   grepl(paste(cond2, ".*", cond1, sep=""), pairwise_results$contrast), ]
    
    if (nrow(pairwise_row) > 0) {
      pairwise_table <- rbind(pairwise_table, data.frame(
        Comparison = paste("Condition", cond1, "vs", cond2),
        Mean_Diff = round(abs(pairwise_row$estimate), 3),
        SE = round(pairwise_row$SE, 3),
        p_adj = ifelse(pairwise_row$p.value < 0.001, "<0.001", round(pairwise_row$p.value, 3)),
        BF10 = round(exp(bf_result@bayesFactor$bf), 3)
      ))
    }
  }
}

print(pairwise_table, row.names = FALSE)
```


## comparing short and long

```{r}
nonVeri_abs.agg <- rbind(mutate(nonVeri_abs.short_agg, exp = "exp 2a"), mutate(nonVeri_abs.long_agg, exp = "exp 2b")) %>% mutate(exp = factor(exp))

# Run t-test
t_result <- t.test(cs.abs ~ exp, data = nonVeri_abs.agg, alternative = "two.sided")
print(t_result)

# Calculate Cohen's d
cohens_d <- cohen.d(cs.abs ~ exp, data = nonVeri_abs.agg)
print(cohens_d)

# Bayes Factor
bf_result <- ttestBF(formula = cs.abs ~ exp, data = nonVeri_abs.agg, alternative = "two.sided")
print(bf_result)
```